---
title: "Spending Habits"
author: "Christopher Eshleman"
date: "5/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Spending habits
I'm going to try and understand my spending habits through exploratory work. 
A sidecar goal is to find ways to reduce personal spending by 5 percent over the next two years. But first things first. 

I went to my bank and credit card websites and figured out how to download spending histories. I could go way back with my checking account / debit card, but only four months or so for my two credit cards. 

```{r}

cat("\014") # clear the console
rm(list=ls())
dev.off() 

check = read.csv("~/Dropbox/Home/Financial/spending/Checking1.csv", header=F) 
names(check) = c("date","amt","na","check_no","desc") 
check$card = "debit" 

cc1 = read.csv("~/Dropbox/Home/Financial/spending/CreditCard4.csv", header=F) 
names(cc1) = c("date","amt","na","check_no","desc") 
cc1$card = "cardWF" 

dat = rbind(check,cc1) 
dat$na = NULL 
dat$check_no = NULL 

``` 

So I'll combine the first file (checking/debit) with my first credit card. 
Then I'll read in the second card, which is from a different institution, and reshape it to match the format from teh first two. Then I'll be able to mash all three together. 

```{r}
cc2 = read.csv("~/Dropbox/Home/Financial/spending/bk_download.csv", header=F) 
cc2$card = "cardUS" 
cc2 = cc2[, c(3,7,5,8)] 
names(cc2) = c("date","amt","desc","card") 

dat = rbind(dat,cc2) 

rm(cc1,cc2,check) 

dat$date = as.Date(dat$date, format="%m/%d/%Y")  #recognizes this variable as a date. 
```

Since only one of them includes dates back to the beginning of 2018, I'll chop off everything before February of this year. That'll give me three and a half months to get started with. 

```{r}
dat = subset(dat,dat$date>"2019-01-31") 
```

## Spending 

I want to know how I spend money by category or type. Groceries versus movie rentals (mostly on Google Play), for example. 

First, some basic stuff. Do I use some cards for bigger purchases than others? 

```{r}
spend = subset(dat,dat$amt<0 & dat$amt>-1200) #Limit to spending, not deposits or payments
spend$amount = -(spend$amt) #Make it positive
summary(spend$amount) 
```

I focused on spending, not deposits or payments. And I knocked out a one-time big-ticket item that shouldn't count as part of my personal spending, as well as monthly rent. 

I have so much spending that falls into the few-dollar range, I'll take the log of spending for graphing purposes. 

```{r, echo=FALSE}
library(grid)
library(ggplot2) 
fill <- "#4271AE"
lines <- "#1F3552"

boxes = ggplot(spend, aes(x = card, y = log(amount))) +
        geom_boxplot(colour = lines, fill = fill,
                     size = 1) 
        #+
        #scale_y_continuous(name = "Purchase amount")#,
                              #breaks = seq(0, 175, 25),
                              #limits=c(0, 175)) 
        #            +
        #scale_x_discrete(name = "Card") +
        #ggtitle("How much I spend vs what I'm using") +
        #theme_bw() +
        #theme(panel.grid.major = element_line(colour = "#d3d3d3"),
        #      panel.grid.minor = element_blank(),
        #      panel.border = element_blank(),
        #      panel.background = element_blank(),
        #      plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        #      text=element_text(family = "Tahoma"),
        #      axis.title = element_text(face="bold"),
        #      axis.text.x = element_text(colour="black", size = 11),
        #      axis.text.y = element_text(colour="black", size = 9),
        #      axis.line = element_line(size=0.5, colour = "black"))
boxes
```

Yep, looks like I've been using my "US" credit card for slightly larger purposes. I wonder why that is ... it's the most recent card I got. 

More to come here ... 

(Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.) 
